---
title: "ESM 244 Asssignment 1: Task 2"
author: "Zoe Duerksen-Salm"
date: "1/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Attach packages
library(tidyverse)
library(janitor)
library(palmerpenguins)
library(here)
library(readxl)
library(dplyr)
library(lubridate)
library(patchwork)
library(ggfortify)
```

A useful descriptive introductory summary (3 - 4 sentences) that helps the audience understand the data (include a citation as necessary) and what you’ll be exploring
All of your organized and well-annotated code (with warnings/messages suppressed) you wrote to wrangle data then run PCA, and to create a professional looking PCA biplot that appears (with a caption) in your knitted HTML
A brief summary (nicely formatted and professionally written bulletpoints are fine) highlighting some major takeaways from your PCA that can be gleaned from the biplot


```{r}
usda_nuts <- read_csv("usda_nutrients.csv") %>% 
  clean_names()

usda_nuts_pca <- usda_nuts %>% 
  select(energy_kcal, fat_g, carb_g, fiber_g, sugar_g, protein_g, vit_a_mcg, vit_c_mg, calcium_mg, iron_mg) %>%
  drop_na() %>% #drop rows containing any NA values
  scale() %>% #scale variables
  prcomp() #run principal components

#see what a list contains [type list$... contents will pop up]
usda_nuts_pca$rotation

# create a biplot of usda_nuts_pca list
# use autoplot fxn (creates the type of plot it assumes you want to make)
autoplot(usda_nuts_pca)+
  theme_minimal()

# BUT WAIT: We want to distinguish these points aesthetically based on other variables too!
# create dataset where observations include what is necessary for the biplot, but still contain other info
usda_nuts_complete <- usda_nuts %>% 
  drop_na(energy_kcal, ends_with("_g"), ends_with("_mg"), ends_with("_mcg"), ends_with("_usrda"))

autoplot(usda_nuts_pca,
         data = usda_nuts_complete,
         colour = "food_group",
         loadings = TRUE,
         loadings.label = TRUE)+
  theme_minimal()+
  theme(legend.position = 'none')

```